Explore Red Wine Quality by Dewi Octavia
========================================================

In this project, a data set of red wine quality will be explored based on its physicochemical properties using the statistical software, R. 
The objective is to find physicochemical properties that distinguish good 
quality wine from lower quality ones. An attempt to build linear model on 
wine quality will also be shown.  

```{r echo=FALSE, message=FALSE, warning=FALSE, packages}
library(ggplot2)
library(gridExtra)
library(GGally)
library(corrplot)
library(memisc)
library(MASS)
library(lattice)
library(knitr)
library(reshape)
library(dplyr)
library(RColorBrewer)
```

```{r echo=FALSE, Load_the_Data}
# Load the Data
setwd("C:/Users/Dewi Octavia/Documents/P4_Explore and Summarize Data/P4_Project Final")
rw <- read.csv("wineQualityReds.csv", sep = ',')
```

## Summary Statistics
Firstly, I would like to understand the data set structure. Summary and str 
functions were used for this purpose. This data set consists of 
1599 observations with 11 physicochemical properties as input variables and 
quality as the output. Wine quality is an ordered and discrete variable, 
the quality ranges from 3.0 to 8.0, with mean and median of 5.6 and 6.0, 
respectively. Each observation is identified in X variable. From the data set 
description, there is a pair of subset (dependant) variables that is 
free sulfur dioxide to total sulfur dioxide. 



# Univariate Plots Section
After the first look of the data set, I will now plot those variables in 
histogram to have a quick glance of the distribution.

```{r echo=FALSE, warning = FALSE, message = FALSE}
p1 <- ggplot(aes(x = fixed.acidity), data = rw) + geom_histogram()
p2 <- ggplot(aes(x = volatile.acidity), data = rw) + geom_histogram()
p3 <- ggplot(aes(x = citric.acid), data = rw) + geom_histogram()
p4 <- ggplot(aes(x = residual.sugar), data = rw) + geom_histogram()
p5 <- ggplot(aes(x = chlorides), data = rw) + geom_histogram()
p6 <- ggplot(aes(x = free.sulfur.dioxide), data = rw) + geom_histogram()
p7 <- ggplot(aes(x = total.sulfur.dioxide), data = rw) + geom_histogram()
p8 <- ggplot(aes(x = density), data = rw) + geom_histogram()
p9 <- ggplot(aes(x = pH), data = rw) + geom_histogram()
p10 <- ggplot(aes(x = sulphates), data = rw) + geom_histogram()
p11 <- ggplot(aes(x = alcohol), data = rw) + geom_histogram()
p12 <- ggplot(aes(x = quality), data = rw) + geom_histogram()

grid.arrange(p1, p2, p3, p4, p5, p6, p7, p8, p9, p10, p11, p12, ncol = 4)
rm(p1, p2, p3, p4, p5, p6, p7, p8, p9, p10, p11, p12)
```


```{r echo=FALSE, warning = FALSE, message = FALSE}
print('=== Stats Summary of pH ===')
summary(rw$pH)

print('=== Stats Summary of Density ===')
summary(rw$density)
```

pH and density appear to be normally distributed. The normal distribution is
confirmed by almost equal mean and median values. Other variables are mostly 
long-tailed with a few outliers. I will replot the long-tailed distributions in 
log scale and compare it to its original plot along with their stats summaries.


```{r echo=FALSE, warning = FALSE, message = FALSE}
compareDist <- function(x1, xlab) {
  grid.arrange(ggplot(aes(x = x1), 
                      data = rw) + 
                 geom_histogram() +
                 xlab(xlab),
               ggplot(aes(x = log10(x1)),
                          data = rw) + 
                 geom_histogram() + 
                 xlab(paste('log(',xlab,')')),
               ncol = 2)
  }
```

```{r echo =FALSE, message=FALSE, warning=FALSE}
compareDist(rw$fixed.acidity, 'fixed.acidity')

print('=== Stats Summary ===')
summary(rw$fixed.acidity)
```

The distribution of fixed acidity in log scale seems to be more normal.

```{r echo =FALSE, message=FALSE, warning=FALSE}
compareDist(rw$volatile.acidity, 'volatile.acidity')

print('=== Stats Summary ===')
summary(rw$volatile.acidity)
```

Distribution of volatile acidity in log scale is also more normal, however it 
still looks slightly skewed.


```{r echo =FALSE, message=FALSE, warning=FALSE}
compareDist(rw$citric.acid, 'citric.acid')

print('=== Stats Summary ===')
summary(rw$citric.acid)

print('Number of zero values')
table(rw$citric.acid == 0)
```

In the original plot, there is suspiciously high count of zero in 
citric acid. I wonder if this is truly zero or simply a 'not available' value. 
A quick check using table function shows that there are 132 observations of zero
values and no NA value in reported citric acid concentration. The citric acid 
concentration could be too low and insignificant hence was reported as zero. 
Replotting citric acid distribution in log scale does not help normalizing the
distribution, it could be due to the high count of zero values mentioned above.



```{r echo =FALSE, message=FALSE, warning=FALSE}
compareDist(rw$residual.sugar, 'residual.sugar')

print('=== Stats Summary ===')
summary(rw$residual.sugar)
```

Residual sugar distribution is more normal in log scale however it is still
long-tailed.


```{r echo =FALSE, message=FALSE, warning=FALSE}
compareDist(rw$chlorides, 'chlorides')

print('=== Stats Summary ===')
summary(rw$chlorides)
```

Distribution of log(chlorides) is more normal than distribution of chlorides.


```{r echo =FALSE, message=FALSE, warning=FALSE}
compareDist(rw$free.sulfur.dioxide, 'free.sulfur.dioxide')

print('=== Stats Summary ===')
summary(rw$free.sulfur.dioxide)
```

Replotting free sulfur dioxide distribution in log scale shows a bimodal 
distribution behaviour.

```{r echo =FALSE, message=FALSE, warning=FALSE}
compareDist(rw$total.sulfur.dioxide, 'total.sulfur.dioxide')

print('=== Stats Summary ===')
summary(rw$total.sulfur.dioxide)
```

Distribution of total sulfur dioxide in log scale is more normal.


```{r echo =FALSE, message=FALSE, warning=FALSE}
compareDist(rw$sulphates, 'sulphates')

print('=== Stats Summary ===')
summary(rw$sulphates)
```

Distribution of log(sulphates) is normal.


```{r echo =FALSE, message=FALSE, warning=FALSE}
compareDist(rw$alcohol, 'alcohol')

print('=== Stats Summary ===')
summary(rw$alcohol)
```

Replotting alcohol in log scale does not normalize the distribution.


```{r echo =FALSE, message=FALSE, warning=FALSE}
compareDist(rw$quality, 'quality')

print('=== Stats Summary ===')
summary(rw$quality)
```

Distribution of quality in log scale is more skewed than its original scale. In
the original plot, we can see that the count of mid range quality (quality of 5
and 6) is considerably higher than others. This might become an issue when 
comparing low and high quality wines.

Wine quality ranges from 3.0 to 8.0 in this data frame. Since I am more inclined
in investigating what makes a higher quality wine, I will add a new variable 
quality.rating to categorise quality values of 3.0-4.0 as 'bad', 5.0-6.0 as 
'average', and 7.0-8.0 as 'good'.

```{r}
rw$quality.rating <- ifelse(rw$quality <5.0, 'bad', 'average')
rw$quality.rating <- ifelse(rw$quality >6.0, 'good',rw$quality.rating)
rw$quality.rating <- ordered(rw$quality.rating, levels = c('bad','average','good'))

rw[1:20, 13:14]
```

```{r fig.height = 3, fig.width = 5,echo=FALSE}
qplot(data = rw, quality.rating)
```

The distribution of quality rating is much higher on the average rating wine as
seen in quality distribution. 
This is likely to cause overplotting therefore I will be comparing only the bad 
and good wines to find distinctive properties that separate these two.



## Univariate Analysis

#### What is/are the main feature(s) of interest in your dataset?
The feature of interest is wine quality.

#### What other features in the dataset do you think will help support your 
#### investigation into your feature(s) of interest?
Based on the data description given, I suspect acids, residual sugar and 
sulfur dioxide  will have effect on the taste hence wine quality.

#### Did you create any new variables from existing variables in the dataset?
Quality.rating variable is created to group the wine quality into three ratings, 
bad, average and good.

#### Of the features you investigated, were there any unusual distributions? 
#### Did you perform any operations on the data to tidy, adjust, or change the 
#### form of the data? If so, why did you do this?
pH and density appear to be normally distributed. Other variables are mostly 
long-tailed with a few outliers. At this stage, I have not change format of 
the data. 





# Bivariate Plots Section
In this section, I will look into comparing two variables and see if there is 
any correlation between them. I have used ggpairs function to spot obvious 
pattern between variables. 

```{r echo=FALSE, message = FALSE, fig.width = 9, fig.height = 6}
ggpairs(subset(rw, select =- X), 
  lower = list(continuous = wrap("points", shape = I('.'))), 
  upper = list(combo = wrap("box", outlier.shape = I('.'))))
```

To better visualise the strong correlation between variables, I will use 
corrplot function
```{r fig.width = 9, fig.height = 6}
# Exclude variable X in correlation calculation
rw.noX <- subset(rw, select =- X)

rw.corr <- cor(rw.noX[sapply(rw.noX, is.numeric)])

corrplot(rw.corr, method = "number")
```
The ggpairs and corrplot functions above highlight correlation between two 
variables. I will only be looking into variables with correlation coefficient 
above 0.30 and below -0.30.  Let's put those data wrangling skill to use and 
show those variables in a new data frame!
```{r}
#Re-arrange the correlation table
rw.corr.melt <- melt(rw.corr)

rw.corr.melt <- subset(rw.corr.melt, 
                       value <=-0.30 | value >=0.30) 

rw.corr.melt <- subset(rw.corr.melt, 
                       value != 1.0) 

rw.corr.melt <- arrange(rw.corr.melt, 
                        desc(value))


# delete the repeated correlation
rw.corr.melt <- rw.corr.melt[-seq(2, 
                                  nrow(rw.corr.melt), 
                                  by =2),]

rw.corr.melt$value <- round(rw.corr.melt$value, 3)

rw.corr.melt
```

Now we have the list of variable pairs, let's plot them!
```{r fixed acidity vs citric acid, echo=FALSE}
ggplot(aes(x = citric.acid, 
           y = fixed.acidity), 
       data = rw) +
  geom_jitter(alpha = 1/5, 
              color = 'blue') + 
  geom_smooth(method = lm, 
              linetype = 2, 
              color = 'red') +
  geom_text(x = 0.85, 
            y = 14, 
            label = paste("italic(r) == ", round(cor(rw$citric.acid,rw$fixed.acidity),2)),
            parse = TRUE)

```

Fixed acidity has a strong positive correlation to citric acid. In the 
description of data attributes, fixed acidity is defined as 'most acids involved
with wine or fixed or nonvolatile (do not evaporate easily)'. I am wondering if
this means citric acid is part of fixed acidity. If it is, other variables that 
correlates well to fixed acidity will also show some correlation to citric acid. 
A quick peek into rw.corr.melt data frame seems to have proven this finding. 
I will discuss more about this as I plot the rest of the graphs.


```{r density vs fixed acidity, echo=FALSE}
ggplot(aes(x = density, 
           y = fixed.acidity), 
       data = rw) +
  geom_jitter(alpha = 1/5, 
              color = 'blue') + 
  geom_smooth(method = lm, 
              linetype = 2, 
              color = 'red') +
  geom_text(x = 1.0025, 
            y = 14, 
            label = paste("italic(r) == ", round(cor(rw$density,rw$fixed.acidity),2)), 
            parse = TRUE)

```

The plot above shows strong positive correlation between fixed acidity and 
density. If our previous suspicion is true, we will also see some correlation 
between density and citric acid.

```{r density vs citric acid, echo=FALSE}
ggplot(aes(x = density, 
           y = citric.acid), 
       data = rw) +
  geom_jitter(alpha = 1/5, 
              color = 'blue') + 
  geom_smooth(method = lm, 
              linetype = 2, 
              color = 'red') +
  geom_text(x = 1.0025, 
            y = 0.85, 
            label = paste("italic(r) == ", round(cor(rw$density,rw$citric.acid),2)), 
            parse = TRUE)
```

The citric acid is indeed correlated to density. Even though the correlation 
is not as strong as fixed acidity to density, the linear regression line seems 
to show some linear relationship between the two variables. Now let's find other
variable that is correlated to fixed acidity and compare it to citric acid.

```{r pH vs. citric acid and fixed acidity, echo=FALSE}
grid.arrange(
  ggplot(aes(x = pH, 
             y = fixed.acidity), 
         data = rw) +
  geom_jitter(alpha = 1/5, 
              color = 'blue') + 
  geom_smooth(method = lm, 
              linetype = 2, 
              color = 'red') +
  geom_text(x = 3.7, 
            y = 14, 
            label = paste("italic(r) == ", round(cor(rw$pH,rw$fixed.acidity),2)), 
            parse = TRUE),
  ggplot(aes(x = pH, 
             y = citric.acid), 
         data = rw) +
  geom_jitter(alpha = 1/5, 
              color = 'blue') + 
  geom_smooth(method = lm, 
              linetype = 2, 
              color = 'red') +
  geom_text(x = 3.7, 
            y = 0.85, 
            label = paste("italic(r) == ", round(cor(rw$pH,rw$citric.acid),2)), 
            parse = TRUE),
  ncol =2)
```

pH is very well correlated to both fixed acidity and citric acid. On second
thought, the strong correlation of pH and density to both acids could just be
common physical properties of acids. The strong correlation we saw in citric
acid and fixed (tartaric) acid could be the result of both acids being 
predominant fixed acids found in wine grapes [Nierman (2004)](http://waterhouse.ucdavis.edu/whats-in-wine/fixed-acidity).

```{r total sulfur dioxide vs free sulfur dioxide, echo=FALSE}
ggplot(aes(x = total.sulfur.dioxide, 
           y = free.sulfur.dioxide), 
       data = rw) +
  geom_jitter(alpha = 1/5, 
              color = 'blue') + 
  geom_smooth(method = lm, 
              linetype = 2, 
              color = 'red') +
  geom_text(x = 250, 
            y = 70, 
            label = paste("italic(r) == ",
                          round(cor(rw$total.sulfur.dioxide,rw$free.sulfur.dioxide),2)), 
            parse = TRUE)
```

From the plot, we can see an clear relationship between free sulfir dioxide and
total sulfur dioxide. This can confirm free sulfur dioxide being subset of total
sulfur dioxide.


```{r alcohol vs quality, echo=FALSE}
ggplot(aes(x = quality, 
           y = alcohol), 
       data = rw) +
  geom_jitter(alpha = 1/5, 
              color = 'blue') + 
  geom_smooth(method = lm, 
              linetype = 2, 
              color = 'red') +
  geom_text(x = 7.5, 
            y = 14.5, 
            label = paste("italic(r) == ", round(cor(rw$alcohol,rw$quality),2)), 
            parse = TRUE)
```

This plot is rather interesting. There is overcrowding in quality of 5 and 6 
due to higher number of mid range quality wine in data set. However if we 
compare the low quality (3-4) to high quality (7-8), there is a trend of 
increasing alcohol content from low to high wine quality. 


```{r sulphates vs chlorides, echo=FALSE}
ggplot(aes(x = chlorides, 
           y = sulphates), 
       data = rw) +
  geom_jitter(alpha = 1/5, 
              color = 'blue') + 
  geom_smooth(method = lm, 
              linetype = 2, 
              color = 'red') +
  geom_text(x = 0.45, 
            y = 1.7, 
            label = paste("italic(r) == ", round(cor(rw$sulphates,rw$chlorides),2)), 
            parse = TRUE)
```

Sulphates and chlorides seems to have some correlation however it is rather poor.


```{r density vs residual sugar, echo=FALSE}
ggplot(aes(x = residual.sugar, 
           y = density), 
       data = rw) +
  geom_jitter(alpha = 1/5, 
              color = 'blue') + 
  geom_smooth(method = lm, 
              linetype = 2, 
              color = 'red') +
  geom_text(x = 14, 
            y = 1, 
            label = paste("italic(r) == ", round(cor(rw$density,rw$residual.sugar),2)),
            parse = TRUE)
```

As sugar solution is denser than water, it is expected to see increasing density
as residual sugar concentration increases. The plot above shows rather weak 
correlation between these two variables. 


```{r sulphates vs citric acid, echo=FALSE}
ggplot(aes(x = sulphates, 
           y = citric.acid), 
       data = rw) +
  geom_jitter(alpha = 1/5, 
              color = 'blue') + 
  geom_smooth(method = lm, 
              linetype = 2, 
              color = 'red') +
  geom_text(x = 1.75, 
            y = 0.9, 
            label = paste("italic(r) == ", round(cor(rw$sulphates,rw$citric.acid),2)), 
            parse = TRUE)
```

There is a weak positive correlation between citric acid and sulphates.


```{r pH vs density, echo=FALSE}
ggplot(aes(x = pH, 
           y = density), 
       data = rw) +
  geom_jitter(alpha = 1/5, 
              color = 'blue') + 
  geom_smooth(method = lm, 
              linetype = 2, 
              color = 'red') +
  geom_text(x = 1.75, 
            y = 0.9, 
            label = paste("italic(r) == ", round(cor(rw$pH,rw$density),2)), 
            parse = TRUE)
```

This plot shows red wine with lower pH tends to have higher density.


```{r volatile acidity vs quality, echo=FALSE}
ggplot(aes(x = quality, 
           y = volatile.acidity), 
       data = rw) +
  geom_jitter(alpha = 1/5, 
              color = 'blue') + 
  geom_smooth(method = lm, 
              linetype = 2, 
              color = 'red') +
  geom_text(x = 7, 
            y = 1.4, 
            label = paste("italic(r) == ", round(cor(rw$volatile.acidity,rw$quality),2)),
            parse = TRUE)
```

The plot above shows negative correlation between volatile acidity and wine 
quality. High quality red Wines have lower volatile (acetic) acid.


```{r alcohol vs density, echo=FALSE}
ggplot(aes(x = alcohol, 
           y = density), 
       data = rw) +
  geom_jitter(alpha = 1/5, 
              color = 'blue') + 
  geom_smooth(method = lm, 
              linetype = 2, 
              color = 'red') +
  geom_text(x = 1.75, 
            y = 0.9, 
            label = paste("italic(r) == ", round(cor(rw$alcohol,rw$density),2)), 
            parse = TRUE)
```

It is expected that decreasing density as alcohol content increases in the plot.
In fermentation process, sugar is turned into alcohol. The more alcohol produced,
the less sugar remains hence lower density.


```{r volatile acidity vs citric acid, echo=FALSE}
ggplot(aes(x = citric.acid, 
           y = volatile.acidity), 
       data = rw) +
  geom_jitter(alpha = 1/5, 
              color = 'blue') + 
  geom_smooth(method = lm, 
              linetype = 2, 
              color = 'red') +
  geom_text(x = 0.85, 
            y = 1.4, 
            label = paste("italic(r) == ", round(cor(rw$citric.acid,rw$volatile.acidity),2)), 
            parse = TRUE)

```

Volatile acidity shows strong negative correlation to citric acid. The strong 
correlation can be explained by subsequent conversion of citric acid to acetic 
acid in a wine making practice involving malo-lactic bacterium [Shimazu et al. (1985)](http://www.tandfonline.com/doi/abs/10.1080/00021369.1985.10867041) 




## Bivariate Analysis

#### Talk about some of the relationships you observed in this part of the 
#### investigation. How did the feature(s) of interest vary with other features 
#### in the dataset?
Variables that show correlation to quality are alcohol and volatile acidity. 
Alcohol content has positive correlation to wine quality. On the other hand, 
volatile acidity is negatively correlated to quality.


#### Did you observe any interesting relationships between the other features 
#### (not the main feature(s) of interest)?
<ul>
<li> Three acids variables appear to be correlated. As citric acid increases, 
fixed acidity increases and volatile acidity decreases. Citric acid can be seen 
negatively correlated to volatile (acetic) acid because citric acid is 
subsequently converted to acetic acid in wine making process. Both fixed 
(tartaric) and volatile (acetic) acids are present in wine grape, this could be 
the reason these two variable are positively correlated. </li> 
<li> Among three acids, volatile acidity does not have as strong correlation to 
pH and density compared to fixed acidity and citric acid.</li>
<li> Residual sugar has an expected correlation to density. Since sugar is 
denser than water solution, density increases as residual sugar increases. </li>
<li> Chlorides seems to be somehow correlated to sulphates.</li>
<li> Free sulfur dioxide is strongly correlated to total sulfur dioxide as they
are dependent (subset) pair variables.</li>
<li> Density is strongly correlated to fixed acidity. A quick check in Wikipedia
reveals tartaric acid has higher density than water.</li>
<li> Alcohol content shows negative correlation to density. In fermentation 
process, sugar is converted to alcohol. The more alcohol produced, the less 
sugar remains hence lower density.  </li>
</ul>


#### What was the strongest relationship you found?
pH to fixed acidity has the strongest relationship, which makes sense as pH is 
the scale to measure acidity. This is followed by fixed acidity to citric acid, 
fixed acidity to density and free sulfur dioxide to total sulfur dioxide.




I would also like to see if any particular variable in log scale has stronger 
correlation to quality. The comparison will be shown in a new data frame, 
'df.corr'.
```{r echo =FALSE, warning=FALSE, message=FALSE}
# Create data frame with variables in log scale
rw.noX <- subset(rw.noX, select =- quality.rating)
logrw <- log(rw.noX)

# Since we will be comparing variables in log to original quality data, I will
# replace log(quality) in logrw with quality
logrw$quality <- rw$quality

# Calculate correlation coefficients
rw.corr <- round(cor(rw.noX, rw.noX$quality), digits = 2)
logrw.corr <- round(cor(logrw, logrw$quality), digits = 2)

df.corr <- data.frame(rw.corr, logrw.corr)
names(df.corr) <- c('Correlation to quality', 'Correlation (log scale) to quality')
df.corr
```

The new data frame above shows that transforming sulphates into log scale 
improves its correlation to quality. The same is observed in chlorides however 
its correlation to wine quality is not as strong as sulphates. Sulphates and 
chlorides will be converted to log scale for the rest of analysis.


I will now compare the original and log variables to quality in a graph (to put 
writing user-defined function into practice). The graphs will be plotted 
side-by-side for comparison, with graph on right handside in log scale. The grid
is splitted into three separate plots for better visuality.

```{r echo=FALSE}
corr.r <- function(x,y) {
corr.coeff <- round(cor(x,y), digits =2)
paste("italic(r) == ", corr.coeff)}
```

```{r echo=FALSE}
get.corr.plot <- function(data, x, y, xtxt, ytxt, xlab) {
ggplot(aes(x = x, y = y), data = data) +
geom_jitter(alpha = 1/5, color = 'blue') +
geom_smooth(method = lm, linetype = 2, color = 'red') +
xlab(xlab) + 
ylab('quality') +
geom_text(x = xtxt, y = ytxt, label = corr.r(x,y), parse = TRUE)}
```

```{r warning=FALSE, echo = FALSE, message= FALSE, fig.width=8, fig.height=8}
grid.arrange(get.corr.plot(rw, rw$fixed.acidity, rw$quality,14,7.5,'fixed acidity'),
            get.corr.plot(rw, log(rw$fixed.acidity), rw$quality,2.5,7.5,'fixed acidity'), 
            get.corr.plot(rw, rw$volatile.acidity, rw$quality,1.2,7,'volatile acidity'),
            get.corr.plot(rw, log(rw$volatile.acidity), rw$quality,0.2,7,'volatile acidity'), 
            get.corr.plot(rw, rw$citric.acid, rw$quality,0.8,7,'citric acid'),
            get.corr.plot(rw, log(rw$citric.acid), rw$quality,-4,7,'citric acid'), 
            get.corr.plot(rw, rw$residual.sugar, rw$quality,14,7.5,'residual.sugar'),
            get.corr.plot(rw, log(rw$residual.sugar), rw$quality,2.5,7.5,'residual.sugar'), 
            ncol =2)
```

```{r warning=FALSE, echo = FALSE, message= FALSE, fig.width=8, fig.height=8}
grid.arrange(get.corr.plot(rw, rw$chlorides, rw$quality,0.5,7,'chlorides'),
            get.corr.plot(rw, log(rw$chlorides), rw$quality,-4,7,'chlorides'), 
            get.corr.plot(rw, rw$free.sulfur.dioxide, rw$quality,60,7,'free.sulfur.dioxide'),
            get.corr.plot(rw, log(rw$free.sulfur.dioxide), rw$quality,0.5,7,'free.sulfur.dioxide'), 
            get.corr.plot(rw, rw$total.sulfur.dioxide, rw$quality,200,7.5,'total.sulfur.dioxide'),
            get.corr.plot(rw, log(rw$total.sulfur.dioxide), rw$quality,5,7.5,'total.sulfur.dioxide'), 
            get.corr.plot(rw, rw$density, rw$quality,1.0025,7.5,'density'),
            get.corr.plot(rw, log(rw$density), rw$quality,0.0025,7.5,'density'), 
            ncol = 2)
```

```{r warning=FALSE, echo = FALSE, message= FALSE, fig.width=8, fig.height=8}
grid.arrange(get.corr.plot(rw, rw$pH, rw$quality,3.7,7.5,'pH'),
            get.corr.plot(rw, log(rw$pH), rw$quality,1.35,7.5,'pH'), 
            get.corr.plot(rw, rw$sulphates, rw$quality,1.5,7.5,'sulphates'),
            get.corr.plot(rw, log(rw$sulphates), rw$quality,0.5,7.5,'sulphates'), 
            get.corr.plot(rw, rw$alcohol, rw$quality,14.5,7.5,'alcohol'),
            get.corr.plot(rw, log(rw$alcohol), rw$quality,2.65,7.5,'alcohol'),
            ncol = 2)
```

From the comparison between original and log scale plots, transforming sulphates
and chlorides seems to improve the correlation slightly. This also reflected
when distribution of sulphates and chlorides were replotted in Univariate Plots 
Section. Distribution of log(sulphates) and log(chlorides) appeared more normal 
than their original plots.



# Multivariate Plots Section
In this section, I will mostly focus on variables that are well correlated to 
quality. They are alcohol content, volatile acidity and sulphates.


```{r echo=FALSE}
ggplot(aes(x = alcohol, 
           y = volatile.acidity, 
           color = factor(quality.rating)), 
       data = rw) + 
  geom_point(alpha = 1/5) + 
  geom_smooth(method = lm, 
              fill =NA, 
              linetype = 1, 
              size = 2)+
  scale_color_brewer(type = 'seq', 
                     palette = 'YlOrRd' ) + 
  theme_dark()
```

This agrees with data attributes description that high level of volatile acid 
gives unpleasant, vinegar taste hence low wine quality. Also, good wines have 
higher alcohol content than bad wines.


```{r echo=FALSE, Multivariate_Plots}
ggplot(aes(x = alcohol, 
           y = density, 
           color = factor(quality.rating)), 
       data = rw) + 
  geom_point(alpha = 1/5) + 
  geom_smooth(method = lm, 
              fill =NA, 
              linetype = 1, 
              size = 2)+
  scale_color_brewer(type = 'seq', 
                     palette = 'YlOrRd' ) + 
  theme_dark()
```

Negative correlation between alcohol and density is consistent in all three 
ratings. The plot also shows that while holding density constant, bad rating 
wine has lower alcohol content compared to good rating wine. It is good to see 
change of slope steepness as the wine rating gets better.    



```{r echo=FALSE}
ggplot(aes(x = alcohol, 
           y = sulphates, 
           color = factor(quality.rating)), 
       data = rw) +
  geom_point(alpha = 1/5) + 
  geom_smooth(method = lm, 
              fill =NA, 
              linetype = 1, 
              size = 2)+
  scale_color_brewer(type = 'seq', 
                     palette = 'YlOrRd' ) + 
  theme_dark()
```

This plot shows better wine tends to have higher sulphates concentration. The 
range of sulphate concentration for a certain wine rating seems to be narrow. 


```{r echo=FALSE}
ggplot(aes(x = volatile.acidity, 
           y = pH, 
           color = factor(quality.rating)), 
       data = rw) +
  geom_point(alpha = 1/5) + 
  geom_smooth(method = lm, 
              fill =NA, 
              linetype = 1,
              size = 2)+
  scale_color_brewer(type = 'seq', 
                     palette = 'YlOrRd' ) + 
  theme_dark()
```

Good wines have lower volatile (acetic) acid than bad and average wines. pH does
not seem to affect quality rating. From volatile acidity below 0.5 g/L, we can 
see that better wines have lower pH when holding volatile acidity constant.  


```{r echo=FALSE}
ggplot(aes(x = volatile.acidity , 
           y = citric.acid, 
           color = quality.rating), 
       data = rw) +
  geom_point(alpha = 1/5) + 
  geom_smooth(method = lm, 
              fill =NA, 
              linetype = 1, 
              size = 2)+
  scale_color_brewer(type = 'seq', 
                     palette = 'YlOrRd' ) + 
  theme_dark()
```


According to its correlation coefficient, citric acid is mildly correlated to 
wine quality. However at low volatile acidity (<0.6 g/L), there is a trend of 
better wine comes with higher citric acid concentration.  



## Linear Model
To build the linear predicting model, I will be using variables with highest 
correlation to wine quality.

```{r echo=FALSE}
m1 <- lm(I(quality) ~ I(alcohol), data = rw)
m2 <- update(m1, ~ . + volatile.acidity)
m3 <- update(m2, ~ . + log(sulphates))
m4 <- update(m3, ~ . + citric.acid)
m5 <- update(m4, ~ . + total.sulfur.dioxide)
m6 <- update(m5, ~ . + density)
mtable(m1, m2, m3, m4, m5, m6, sdigits = 3)

df.predict_error <- data.frame(rw$quality, predict(m6, rw) - as.numeric(rw$quality))

names(df.predict_error) <- c("quality","prediction.error")

ggplot(aes(x = factor(quality), 
           y = prediction.error), 
       data = df.predict_error) + 
  geom_boxplot() + 
  geom_jitter(alpha = 1/5, 
              color = 'deepskyblue1') + 
  stat_summary(fun.y = mean, 
               geom = 'point', 
               shape = '*', 
               size = 8, 
               color = 'red') +
  geom_hline(linetype = 1, 
             alpha = 0.3, 
             yintercept = 0, 
             color = 'red') + 
  theme_dark()
```

R-squared from the model is rather low, this could be due to lack of variable 
that shows strong correlation to wine quality. The predicting model seems to fit
better to average rating wine, this may be caused by larger distribution of 
average wine in data set. 


## Multivariate Analysis

#### Talk about some of the relationships you observed in this part of the 
#### investigation. Were there features that strengthened each other in terms of 
#### looking at your feature(s) of interest?
Low volatile acidity combined with high alcohol content and sulphates seem to 
make better wines.

#### Were there any interesting or surprising interactions between features?
Negative correlation between alcohol and density is consistent in all three 
ratings. The plot also shows bad rating wine has lower alcohol content compared 
to good rating wine.


------

# Final Plots and Summary

### Plot One: Density vs. Alcohol Content by Wine Rating
```{r echo=FALSE, Plot_One}
ggplot(aes(x = alcohol, 
           y = density, 
           color = factor(quality.rating)),
       data = rw) + 
  geom_point(alpha = 1/5) + 
  geom_smooth(method = lm, 
              fill =NA, 
              linetype = 1, 
              size = 2)+
  labs(color = "Quality Rating") + 
  xlab(bquote("Alcohol Content (% by Volume)")) + 
  ylab(bquote("Density (g/cm"^"3"*")")) +
  ggtitle("Density vs. Alcohol Content by Wine Quality Rating") +
  scale_color_brewer(type = 'seq', 
                     palette = 'YlOrRd' ) + 
  theme_dark()
```

#### Description One
This plot shows the influence of alcohol content and density to wine rating. 
Negative correlation between alcohol and density is consistent in all three wine
ratings. This negative correlation can be explained by the fermentation process 
in wine making. Sugar content is directly proportional to density, higher sugar 
content leads to higher density. In fermentation process, sugar is converted to 
alcohol. The more alcohol produced, the less sugar remains hence lower density. 
The change of slope steepness as the wine rating gets better is shown as 
expected. The plot also shows that while holding density constant, bad rating 
wine has lower alcohol content compared to good rating wine. The average rating 
wine data is ignored in inferring relationship between variables due to 
significantly higher number of average rating data. 


### Plot Two: Sulphates vs. Alcohol Content by Wine Rating
```{r echo=FALSE, Plot_Two}
ggplot(aes(x = alcohol, 
           y = sulphates, 
           color = factor(quality.rating), 
           labs(color = 'Quality Rating')), 
       data = rw) +
  geom_point(alpha = 1/5) + 
  geom_smooth(method = lm, 
              fill =NA, 
              linetype = 1, 
              size = 2)+
  labs(color = "Quality Rating") + 
  xlab(bquote("Alcohol Content (% by Volume)")) +
  ylab(bquote("Sulphates (g/dm"^"3"*")")) +
  ggtitle("Sulphates vs. Alcohol Content by Wine Quality Rating") +
  scale_color_brewer(type = 'seq', 
                     palette = 'YlOrRd' ) + 
  theme_dark()
```

#### Description Two
This chart reveals influence of alcohol and sulphates concentration to red wine 
rating. It shows that better wines tend to have higher sulphates and alcohol 
concentrations. The range of sulphate concentration for a certain wine rating 
seems to be small.  


### Plot Three: Linear Predicting Model
```{r echo=FALSE, Plot_Three}
ggplot(aes(x = factor(quality), 
           y = prediction.error), 
       data = df.predict_error) + 
  geom_boxplot() + 
  geom_jitter(alpha = 1/5, 
              color = 'deepskyblue1') + 
  stat_summary(fun.y = mean, 
               geom = 'point', 
               shape = '*', 
               size = 8, 
               color = 'red') +
  geom_hline(linetype = 1, 
             alpha = 0.3, 
             yintercept = 0, 
             color = 'red') +
  xlab("Quality") +
  ylab("Model Prediction Error") +
  ggtitle("Linear Model Prediction Error vs. Red Wine Quality") + 
  theme_dark()
```

#### Description Three
With R-squared score of 35.3%, the linear predicting model does not help explain
the variance in wine quality. Although the model generated shows better 
correlation to average rating wine, this could be due to high number of average 
rating wine in data set or missing other key properties that better predict wine
quality.

------

# Reflection
In this project, I was able to examine relationship between physicochemical 
properties and identify the key variables that determine red wine quality, which
are alcohol content and volatile acidity. Some interesting findings of 
relationship between variables was made sensible using scientific explanation 
such as relationship between alcohol content, residual sugar and density in wine
making process. Data wrangling skill was put into practice in this project for 
rearranging data into a suitable format. Lack of variable that shows strong 
correlation to wine quality and high distribution of average rating wine proved 
to be problematic in performing analysis. It was hard to tell if a true 
correlation was present. This also shows limitation in generating an accurate 
predicting model. For future data exploration, it will be interesting to apply 
different approach in building the algorithm and look into evaluations made by 
each wine experts as wine tasting is subject to individual preferences.  